Visualizing Independence Using Extended Association Plots

### 2005-07-12 trunk/vcd/paper/JCGS.Rnw
started changes for JCGS paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@392 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

data(Arthritis)
art <- xtabs(~ Improved + Treatment, data = Arthritis, subset = Sex == "Female")

data(Employment)
emp <- Employment
emp1 <- co.table(Employment, "LayoffCause")[["Replaced"]]
@

<<auxfunctions,echo=FALSE,results=hide>>=
pmos.Friendly <- function(x, ...) grid.mosaicplot(x, panel = TRUE, legend = FALSE,
  labels = FALSE, space = 0.02, gp = gp.Friendly, ...)
pmos.max <- function(x, ...) grid.mosaicplot(x, panel = TRUE, legend = FALSE,
  labels = FALSE, space = 0.02, gp = gp.max, ...)

hue.slice <- function(hue, grid.n = 100, type = c("HCL", "HSV"), plot = TRUE, correct = FALSE)
{
  type <- match.arg(type)
  if(type == "HCL") {
    chroma = seq(0, 100, length = grid.n)
    luminance = seq(0, 100, length = grid.n)
    nc <- length(chroma)
    nl <- length(luminance)
    color.slice <- outer(chroma, luminance, function(y, x) hcl(hue, x, y, verbose = FALSE, correct = correct))
    xlab <- "chroma"
    ylab <- "luminance"
    main <- paste("hue =", round(hue, digits = 0))
  } else {
    chroma = seq(0, 1, length = grid.n)
    luminance = seq(0, 1, length = grid.n)
    nc <- length(chroma)
    nl <- length(luminance)
    color.slice <- outer(chroma, luminance, function(y, x) hsv(hue, x, y))
    xlab <- "saturation"
    ylab <- "value"
    main <- paste("hue =", round(hue, digits = 3))
  }
  if(plot) {
    plot(0.5, 0.5, xlim = range(chroma), ylim = range(luminance), type = "n", axes = FALSE,
         xlab = xlab, ylab = ylab, yaxs = "i", xaxs = "i", main = main)
    for(i in 1:(nc-1)) {
      rect(chroma[i], luminance[-nl], chroma[i] + 100/(nc-1), luminance[-1], border = color.slice[,i+1], col = color.slice[,i+1])
    }
    axis(1)
    axis(2)
    box()
  }
  colnames(color.slice) <- chroma
  rownames(color.slice) <- luminance
  attr(color.slice, "type") <- type
  class(color.slice) <- "slice"
  invisible(color.slice)
}

@@

\begin{abstract}
Story starts out not from exploratory analysis with \proglang{R}
but from the problem of visualizing independence. Several
exploratory approaches but not combined with significance
testing.

Paper could be structured in the following way
\begin{itemize}
  \item Introduction
  \item Tests
  \item Data sets
  \item Visualization Techniques
    \begin{itemize}
      \item Mosaic displays
      \item Association plots
      \item Residuals-based shading
    \end{itemize}
  \item Extensions
    \begin{itemize}
      \item max-shading
      \item HCL colors
      \item conditional independence (coplots)
      \item \textit{not the implementation, just mention in summary}
    \end{itemize}
  \item Summary
\end{itemize}

The \proglang{R} side then has to be described in another paper,
probably for the Journal of Statistical Software.\\[0.3cm]
\textbf{Key Words:} Association plots; Conditional inference; 
  Contingency tables; HCL colors; HSV colors; Mosaic plots; Permutation tests.
\end{abstract}

@@

Statistical models built for the analysis of multivariate data
quickly become complex with increasing dimensionality.
One idea of visualization techniques is to
use the human visual system to detect structures in the data that
possibly are not obvious from solely numeric output (e.g., test
statistics).
The \textsf{R} package \texttt{vcd}---inspired by the Book ``Visualizing Categorical
Data'' \citep{vcd:Friendly:2000}---includes methods for the (mostly
graphical) exploration of categorical data, such as:
\begin{itemize}
 \item fitting and graphing of discrete distributions,
 \item plots and tests for independence and symmetry problems,
 \item visualization techniques for log-linear models.
\end{itemize}
Here, we focus on the visualization of independence,
in particular in 2-way tables.

\section{Tests for Independence in 2-way Tables} \label{sec:tests}

@@

\section{Visualization Techniques for 2-way Tables} \label{sec:visualization}

@@

\subsection{Association Plots}

Association plots \citep{vcd:Cohen:1980} visualize the table of Pearson residuals:
each cell is represented by a rectangle that has (signed) height
proportional to the corresponding Pearson residual $r_{ij}$ and width proportional to the
square root of the expected counts $\sqrt{\hat n_{ij}}$. Thus, the area is proportional to
the raw residuals $n_{ij} - \hat n_{ij}$. The sign of the residual is redundantly
coded by the rectangle's color and its position relative to the baseline.

@@

\subsection{Mosaic Plots}

Mosaic plots can be seen as an extension of grouped
bar charts where width and height of the bars show the relative
frequencies of the two variables: a mosaic plot simply
consists of a collection of tiles whose
sizes are proportional to the observed cell frequencies
(see Figure \ref{fig:goals-mosaic-classic}).

@@

Figure \ref{fig:goals-mosaic-classic} shows the Friendly mosaic plot for
the same Bundesliga data as in Figure \ref{fig:goals-assoc-classic}.
In this plot, the residuals
are visualized by the color shading of the cells. Again, the cells for
the 1--0 and 1--1 results are highlighted and furthermore, there seem to be
too many 6--3 results than would be plausible under independence. Whereas
the association plot in Figure \ref{fig:goals-assoc-classic} contains
no information about the result of a test for independence, the shading
in this plot conveys the impression that there might be
significant dependence (which would indicate that there
is not only {\em no} home field advantage in the German Bundesliga in 1995/6 but even
an advantage for the away team). However, it is unclear whether there
is evidence for this or not.

\section{Extensions} \label{sec:extensions}

In this section, we suggest some improvements and extensions which can be
applied to both visualization techniques introduced in Section
\ref{sec:visualization}. First, the visual enhancements
\begin{itemize}
\item combining visualization of and testing for independence
\item improved color schemes
\end{itemize}
are addressed. Second, the implementation enhancements
\begin{itemize}
\item more modular implementation using \texttt{grid}
\item more flexible plotting parameter specification
\end{itemize}
are described. None of these should be thought of as finished work
but rather as a collection of ideas which still need thorough
refinement and proper implementation.

\subsection{Visualization Enhancements}

@@

\subsection{Implementation Enhancements}

The current implementation of association and mosaic plots in \textsf{R} suffers
from two main disadvantages: First, it is not easy to recycle the
plots in conditioning plots or pairs plots (like mosaic matrices) as
they have been implemented using \textsf{R}'s base graphics engine where in
general plotting to relative coordinates is not supported.
The new implementation was written from scratch in \texttt{grid} offering
much more versatility amongst some minor advantages and convenient
improvements. Second, the graphics parameters of the rectangles in
association and mosaic plots (like in almost all standard \textsf{R}
plots), such as color and line type, cannot be specified for each cell
by the user. To overcome this the current implementation in \texttt{vcd}
allows the user to specify either arrays of graphics parameters
of the same dimensionality as the object being plotted or a function
which computes these graphics parameters based on the original table
and its Pearson residuals. Functions are provided for the shading schemes
described in the previous section. Without this feature we would not be
able to specify the shading schemes described in the previous section
in such a flexible way.

\section{Multi-way Tables}

Independence problems do not only occur in 2-way tables, although that
is an important special case, but they are also important in tables of
higher dimensionality and can follow much more complex patterns.
These are again defined based on the underlying table of theoretical cell probabilities
$(\pi_{ij\dots})$ with more than two dimensions. Models of interest include the
null hypotheses of:
\begin{itemize}
 \item total independence:
  $\pi_{ijk\dots} = \pi_{i++\dots}\pi_{+j+\dots}\pi_{++k\dots}\dots$
 \item conditional independence:
  $\pi_{ijk\dots} = \pi_{i|k\dots}\pi_{j|k\dots}$
 \item joint independence:
  $\pi_{ijk\dots} = \pi_{ij+\dots}\pi_{++k\dots}$
\end{itemize}
Classical non-graphical methods for these problems include the $\chi^2$
test, Fisher's exact test, the Cochran-Mantel-Haenzel test (for $2 \times 2 \times
K$-tables), and the analysis of log-linear models for more
complex settings.

@@

\section{Conclusions}
We suggest a set of enhancements for visualizing the independence
problem in 2-way tables with some outlook to multi-way tables. The
extensions aim at improving the visualization by displaying both
the size of the residuals and the significance of a test for
independence and by using better color schemes. Furthermore, a new
implementation is outlined based on the graphics package \texttt{grid}
which provides more modular design and more flexible specification
of graphical parameters.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[htbp]
\begin{center}
<<goals-assoc-classic,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
grid.assocplot(emp1, gp = gp.binary)
@
<<goals-mosaic-classic,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaicplot(t(emp1), shade = TRUE, main = "")
@
\caption{Classic association and mosaic plot for the ``Replaced'' subset of the Employment data.}
\label{fig:classic}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<emp1-Friendly,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
grid.mosaicplot(emp1, gp = gp.Friendly)
@
<<emp1-max,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
grid.mosaicplot(emp1, gp = gp.max)
@
\caption{Mosaic plots for the ``Replaced'' subset of the Employment data.}
\label{fig:emp1}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<art-Friendly,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
grid.mosaicplot(art, gp = gp.Friendly)
@
<<art-max,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
grid.mosaicplot(art, gp = gp.max)
@
\caption{Mosaic plots for the Arthritis data.}
\label{fig:art}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[htbp]
\begin{center}
<<Friendly-hue,fig=TRUE,echo=FALSE,height=4.8,width=8>>=
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(dummy[-101] + 1/200, 1, 1)
rect(dummy[-101], 0.6, dummy[-1], 1, border = dummy.col, col = dummy.col)
rect(0, 0.6, 1, 1)
axis(3)
mtext("hue", line = 2, cex = 1.5)
text(1, 0.45, "saturation = 1\nvalue = 1", pos = 2, cex = 1.2)
lines(c(0.1, 0), c(0.3, 0.6), lty = 2)
lines(c(2/3, 2/3), c(0.3, 0.6), lty = 2)
rect(0, 0.1, 0.2, 0.3, col = hsv(0, 1, 1))
rect(2/3 - 0.1, 0.1, 2/3 + 0.1, 0.3, col = hsv(2/3, 1, 1))
text(0.1, 0.1, expression(r[ij] < 0), pos = 1, cex = 1.5)
text(2/3, 0.1, expression(r[ij] > 0), pos = 1, cex = 1.5)
@

<<Friendly-sat,fig=TRUE,echo=FALSE,height=4.8,width=8>>=
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(2/3, dummy[-101] + 1/200, 1)
rect(dummy[-101], 0.8, dummy[-1], 1, border = dummy.col, col = dummy.col)
dummy.col <- hsv(0, dummy[-101] + 1/200, 1)
rect(dummy[-101], 0.6, dummy[-1], 0.8, border = dummy.col, col = dummy.col)
rect(0, 0.8, 1, 1)
rect(0, 0.6, 1, 0.8)
axis(3)
mtext("saturation", line = 2, cex = 1.5)
mtext(c("h = 0", "h = 2/3"), side = 2, at = c(0.7, 0.9), las = 1, cex = 1.2)
text(0.8, 0.5, "value = 1", pos = 2, cex = 1.2)
lines(c(0.1, 0), c(0.3, 0.6), lty = 2)
lines(c(0.5, 0.5), c(0.3, 0.6), lty = 2)
lines(c(0.9, 1), c(0.3, 0.6), lty = 2)
rect(0, 0.2, 0.2, 0.3, col = hsv(2/3, 0, 1))
rect(0, 0.1, 0.2, 0.2, col = hsv(0, 0, 1))
rect(0.4, 0.2, 0.6, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.4, 0.1, 0.6, 0.2, col = hsv(0, 0.5, 1))
rect(0.8, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.8, 0.1, 1, 0.2, col = hsv(0, 1, 1))
text(0.1, 0.1, expression(abs(r[ij]) < 2), pos = 1, cex = 1.5)
text(0.5, 0.1, expression(2 < abs(r[ij]) < 4), pos = 1, cex = 1.5)
text(0.9, 0.1, expression(abs(r[ij]) > 4), pos = 1, cex = 1.5)
@

<<Friendly-val,fig=TRUE,echo=FALSE,height=4.8,width=8>>=
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(2/3, 1, dummy[-101] + 1/200)
rect(dummy[-101], 0.8, dummy[-1], 1, border = dummy.col, col = dummy.col)
dummy.col <- hsv(0, 1, dummy[-101] + 1/200)
rect(dummy[-101], 0.6, dummy[-1], 0.8, border = dummy.col, col = dummy.col)
rect(0, 0.8, 1, 1)
rect(0, 0.6, 1, 0.8)
axis(3)
mtext("value", line = 2, cex = 1.5)
mtext(c("h = 0", "h = 2/3"), side = 2, at = c(0.7, 0.9), las = 1, cex = 1.2)
text(0, 0.5, paste("saturation =", 1), pos = 4, cex = 1.2)
lines(c(0.85, 1), c(0.3, 0.6), lty = 2)
rect(0.7, 0.2, 0.8, 0.3, col = hsv(2/3, 0, 1))
rect(0.8, 0.2, 0.9, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.9, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.7, 0.1, 0.8, 0.2, col = hsv(0, 0, 1))
rect(0.8, 0.1, 0.9, 0.2, col = hsv(0, 0.5, 1))
rect(0.9, 0.1, 1, 0.2, col = hsv(0, 1, 1))
lines(c(0.5, 0.5), c(0.3, 0.6), lty = 2)
lines(c(0.85, 1), c(0.3, 0.6), lty = 2)
rect(0.35, 0.2, 0.45, 0.3, col = hsv(2/3, 0, 0.5))
rect(0.45, 0.2, 0.55, 0.3, col = hsv(2/3, 0.5, 0.5))
rect(0.55, 0.2, 0.65, 0.3, col = hsv(2/3, 1, 0.5))
rect(0.35, 0.1, 0.45, 0.2, col = hsv(0, 0, 0.5))
rect(0.45, 0.1, 0.55, 0.2, col = hsv(0, 0.5, 0.5))
rect(0.55, 0.1, 0.65, 0.2, col = hsv(0, 1, 0.5))
rect(0.7, 0.2, 0.8, 0.3, col = hsv(2/3, 0, 1))
rect(0.8, 0.2, 0.9, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.9, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.7, 0.1, 0.8, 0.2, col = hsv(0, 0, 1))
rect(0.8, 0.1, 0.9, 0.2, col = hsv(0, 0.5, 1))
rect(0.9, 0.1, 1, 0.2, col = hsv(0, 1, 1))
text(0.5, 0.1, "non-significant", pos = 1, cex = 1.5)
text(0.85, 0.1, "significant", pos = 1, cex = 1.5)
@

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[htbp]
\begin{center}
<<UCB-mosaic1,fig=TRUE,echo=FALSE,height=6,width=8>>=
fm <- loglm(~(EmploymentStatus+EmploymentLength)*LayoffCause, data = emp, fitted = TRUE)
my.gp <- gp.HCLshading(emp, residuals(fm, type = "pearson"))
my.gp$p.value <- summary(fm)$tests[2,3]
my.gp$lty <- c(1,1)
grid.mosaicplot(emp, gp = my.gp, margin = ~(EmploymentStatus+EmploymentLength)*LayoffCause,
  space = c(0.1, 0.01, 0), dir = "v")
@
\caption{Employment data with mosaic display.}
\label{fig:emp-mosaic1}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<UCB-mosaic2,fig=TRUE,echo=FALSE,height=4,width=8>>=
tabplot(emp, panel = pmos.max)
@
\caption{Employment data with conditioned mosaic display.}
\label{fig:emp-mosaic2}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<UCB-assoc,fig=TRUE,echo=FALSE,height=4,width=8>>=
tabplot(emp)
@
\caption{Employment data with conditioned association plot.}
\label{fig:emp-assoc}
\end{center}
\end{figure}

@@

>>>>

data(Arthritis)
art <- xtabs(~ Treatment + Improved, data = Arthritis, subset = Sex == "Female")

data(Employment)
emp <- aperm(Employment)
emp1 <- co_table(Employment, "LayoffCause")[["Replaced"]]
@

<<auxfunctions,echo=FALSE,results=hide>>=
## pmos.Friendly <- function(x, ...) grid.mosaicplot(x, panel = TRUE, legend = FALSE,
##   labels = FALSE, space = 0.02, gp = gp.Friendly, ...)
## pmos.max <- function(x, ...) grid.mosaicplot(x, panel = TRUE, legend = FALSE,
##   labels = FALSE, space = 0.02, gp = gp.max, ...)

hue.slice <- function(hue, grid.n = 100, type = c("HCL", "HSV"), plot = TRUE, fixup = FALSE)
{
  type <- match.arg(type)
  if(type == "HCL") {
    chroma = seq(0, 100, length = grid.n)
    luminance = seq(0, 100, length = grid.n)
    nc <- length(chroma)
    nl <- length(luminance)
    color.slice <- outer(chroma, luminance, function(y, x) hcl(hue, x, y, fixup = fixup))
    xlab <- "chroma"
    ylab <- "luminance"
    main <- paste("hue =", round(hue, digits = 0))
  } else {
    chroma = seq(0, 1, length = grid.n)
    luminance = seq(0, 1, length = grid.n)
    nc <- length(chroma)
    nl <- length(luminance)
    color.slice <- outer(chroma, luminance, function(y, x) hsv(hue, x, y))
    xlab <- "saturation"
    ylab <- "value"
    main <- paste("hue =", round(hue, digits = 3))
  }
  if(plot) {
    plot(0.5, 0.5, xlim = range(chroma), ylim = range(luminance), type = "n", axes = FALSE,
         xlab = xlab, ylab = ylab, yaxs = "i", xaxs = "i", main = main)
    for(i in 1:(nc-1)) {
      rect(chroma[i], luminance[-nl], chroma[i] + 100/(nc-1), luminance[-1], border = color.slice[,i+1], col = color.slice[,i+1])
    }
    axis(1)
    axis(2)
    box()
  }
  colnames(color.slice) <- chroma
  rownames(color.slice) <- luminance
  attr(color.slice, "type") <- type
  class(color.slice) <- "slice"
  invisible(color.slice)
}

@@

\begin{abstract}

Visualizing independence by association and mosaic plots.

Enhancements: diverging color palette based on HCL color space
combined with visualization of the result of a significance test.

Extensions to visualizing conditional independence.\\[0.3cm]

\textbf{Key Words:} Association plots; Conditional inference; 
  Contingency tables; HCL colors; HSV colors; Mosaic plots; Permutation tests.
\end{abstract}

@@

independence in 2-way tables

association and mosaic displays

colors and color spaces

\section{Independence in 2-way Tables}
\label{sec:2way}

\subsection{Tests}
\label{sec:tests}

@@

\section{Visualizations} \label{sec:visualization}

ONLY SHAPE! not color

@@

\textbf{Association Plots}
\citep{vcd:Cohen:1980} visualize the table of Pearson residuals:
each cell is represented by a rectangle that has (signed) height
proportional to the corresponding Pearson residual $r_{ij}$ and width proportional to the
square root of the expected counts $\sqrt{\hat n_{ij}}$. Thus, the area is proportional to
the raw residuals $n_{ij} - \hat n_{ij}$. The sign of the residual is redundantly
coded by the rectangle's color and its position relative to the baseline.

@@

\subsection{Mosaic Plots}
\citep{vcd:Hartigan+Kleiner:1981} can be seen as an extension of grouped
bar charts where width and height of the bars show the relative
frequencies of the two variables: a mosaic plot simply
consists of a collection of tiles whose
sizes are proportional to the observed cell frequencies
(see Figure \ref{fig:goals-mosaic-classic}).

@@

\section{Residual-based Shadings} \label{sec:shadings}

remarks: Friendly 1994

\subsection{Colors}

\subsection{Significance}

@@

\section{Extensions to Conditional Independence}

Independence problems do not only occur in 2-way tables, although that
is an important special case, but they are also important in tables of
higher dimensionality and can follow much more complex patterns.
These are again defined based on the underlying table of theoretical cell probabilities
$(\pi_{ij\dots})$ with more than two dimensions. Models of interest include the
null hypotheses of conditional independence:
  \[ \pi_{ijk\dots} = \pi_{i|k\dots}\pi_{j|k\dots} \]

Classical non-graphical methods for these problems include the $\chi^2$
test, Fisher's exact test, the Cochran-Mantel-Haenzel test (for $2 \times 2 \times
K$-tables), and the analysis of log-linear models for more
complex settings.

@@

\section{Conclusions}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[htbp]
\begin{center}
<<goals-assoc-classic,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
assoc(emp1)
@
<<goals-mosaic-classic,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(emp1, split_vertical = TRUE)
@
\caption{Classic association and mosaic plot for the ``Replaced'' subset of the Employment data.}
\label{fig:classic}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<art-Friendly,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(art, gp = shading_Friendly)
@
<<art-max,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(art, gp = shading_max)
@
\caption{Mosaic plots for the Arthritis data.}
\label{fig:art}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[htbp]
\begin{center}
<<UCB-mosaic1,fig=TRUE,echo=FALSE,height=6,width=8>>=
## fm <- loglm(~(EmploymentStatus+EmploymentLength)*LayoffCause, data = emp, fitted = TRUE)
## my.gp <- gp.HCLshading(emp, residuals(fm, type = "pearson"))
## my.gp$p.value <- summary(fm)$tests[2,3]
## my.gp$lty <- c(1,1)
## grid.mosaicplot(emp, gp = my.gp, margin = ~(EmploymentStatus+EmploymentLength)*LayoffCause,
##   space = c(0.1, 0.01, 0), dir = "v")
mosaic(emp, expected = ~(EmploymentStatus+EmploymentLength)*LayoffCause)
@
\caption{Employment data with mosaic display.}
\label{fig:emp-mosaic1}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[htbp]
\begin{center}
<<Friendly-hue,fig=TRUE,echo=FALSE,height=4.8,width=8>>
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(dummy[-101] + 1/200, 1, 1)
rect(dummy[-101], 0.6, dummy[-1], 1, border = dummy.col, col = dummy.col)
rect(0, 0.6, 1, 1)
axis(3)
mtext("hue", line = 2, cex = 1.5)
text(1, 0.45, "saturation = 1\nvalue = 1", pos = 2, cex = 1.2)
lines(c(0.1, 0), c(0.3, 0.6), lty = 2)
lines(c(2/3, 2/3), c(0.3, 0.6), lty = 2)
rect(0, 0.1, 0.2, 0.3, col = hsv(0, 1, 1))
rect(2/3 - 0.1, 0.1, 2/3 + 0.1, 0.3, col = hsv(2/3, 1, 1))
text(0.1, 0.1, expression(r[ij] < 0), pos = 1, cex = 1.5)
text(2/3, 0.1, expression(r[ij] > 0), pos = 1, cex = 1.5)
@

<<Friendly-sat,fig=TRUE,echo=FALSE,height=4.8,width=8>>
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(2/3, dummy[-101] + 1/200, 1)
rect(dummy[-101], 0.8, dummy[-1], 1, border = dummy.col, col = dummy.col)
dummy.col <- hsv(0, dummy[-101] + 1/200, 1)
rect(dummy[-101], 0.6, dummy[-1], 0.8, border = dummy.col, col = dummy.col)
rect(0, 0.8, 1, 1)
rect(0, 0.6, 1, 0.8)
axis(3)
mtext("saturation", line = 2, cex = 1.5)
mtext(c("h = 0", "h = 2/3"), side = 2, at = c(0.7, 0.9), las = 1, cex = 1.2)
text(0.8, 0.5, "value = 1", pos = 2, cex = 1.2)
lines(c(0.1, 0), c(0.3, 0.6), lty = 2)
lines(c(0.5, 0.5), c(0.3, 0.6), lty = 2)
lines(c(0.9, 1), c(0.3, 0.6), lty = 2)
rect(0, 0.2, 0.2, 0.3, col = hsv(2/3, 0, 1))
rect(0, 0.1, 0.2, 0.2, col = hsv(0, 0, 1))
rect(0.4, 0.2, 0.6, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.4, 0.1, 0.6, 0.2, col = hsv(0, 0.5, 1))
rect(0.8, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.8, 0.1, 1, 0.2, col = hsv(0, 1, 1))
text(0.1, 0.1, expression(abs(r[ij]) < 2), pos = 1, cex = 1.5)
text(0.5, 0.1, expression((2 < abs(r[ij])) < 4), pos = 1, cex = 1.5)
text(0.9, 0.1, expression(abs(r[ij]) > 4), pos = 1, cex = 1.5)
@

<<Friendly-val,fig=TRUE,echo=FALSE,height=4.8,width=8>>
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(2/3, 1, dummy[-101] + 1/200)
rect(dummy[-101], 0.8, dummy[-1], 1, border = dummy.col, col = dummy.col)
dummy.col <- hsv(0, 1, dummy[-101] + 1/200)
rect(dummy[-101], 0.6, dummy[-1], 0.8, border = dummy.col, col = dummy.col)
rect(0, 0.8, 1, 1)
rect(0, 0.6, 1, 0.8)
axis(3)
mtext("value", line = 2, cex = 1.5)
mtext(c("h = 0", "h = 2/3"), side = 2, at = c(0.7, 0.9), las = 1, cex = 1.2)
text(0, 0.5, paste("saturation =", 1), pos = 4, cex = 1.2)
lines(c(0.85, 1), c(0.3, 0.6), lty = 2)
rect(0.7, 0.2, 0.8, 0.3, col = hsv(2/3, 0, 1))
rect(0.8, 0.2, 0.9, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.9, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.7, 0.1, 0.8, 0.2, col = hsv(0, 0, 1))
rect(0.8, 0.1, 0.9, 0.2, col = hsv(0, 0.5, 1))
rect(0.9, 0.1, 1, 0.2, col = hsv(0, 1, 1))
lines(c(0.5, 0.5), c(0.3, 0.6), lty = 2)
lines(c(0.85, 1), c(0.3, 0.6), lty = 2)
rect(0.35, 0.2, 0.45, 0.3, col = hsv(2/3, 0, 0.5))
rect(0.45, 0.2, 0.55, 0.3, col = hsv(2/3, 0.5, 0.5))
rect(0.55, 0.2, 0.65, 0.3, col = hsv(2/3, 1, 0.5))
rect(0.35, 0.1, 0.45, 0.2, col = hsv(0, 0, 0.5))
rect(0.45, 0.1, 0.55, 0.2, col = hsv(0, 0.5, 0.5))
rect(0.55, 0.1, 0.65, 0.2, col = hsv(0, 1, 0.5))
rect(0.7, 0.2, 0.8, 0.3, col = hsv(2/3, 0, 1))
rect(0.8, 0.2, 0.9, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.9, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.7, 0.1, 0.8, 0.2, col = hsv(0, 0, 1))
rect(0.8, 0.1, 0.9, 0.2, col = hsv(0, 0.5, 1))
rect(0.9, 0.1, 1, 0.2, col = hsv(0, 1, 1))
text(0.5, 0.1, "non-significant", pos = 1, cex = 1.5)
text(0.85, 0.1, "significant", pos = 1, cex = 1.5)
@

\begin{figure}[htbp]
\begin{center}
<<emp1-Friendly,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>
mosaic(emp1, gp = shading_Friendly)
@
<<emp1-max,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>
mosaic(emp1, gp = shading_max)
@
\caption{Mosaic plots for the ``Replaced'' subset of the Employment data.}
\label{fig:emp1}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<UCB-mosaic2,fig=TRUE,echo=FALSE,height=4,width=8>>
tabplot(emp, panel = pmos.max)
@
\caption{Employment data with conditioned mosaic display.}
\label{fig:emp-mosaic2}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
<<UCB-assoc,fig=TRUE,echo=FALSE,height=4,width=8>>
tabplot(emp)
@
\caption{Employment data with conditioned association plot.}
\label{fig:emp-assoc}
\end{center}
\end{figure}
